QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-09-13 paper.tex
Rev_RelatedWorkTable_V0.4

\lettrine{M}{ \,OBILE} edge computing (MEC) \cite{mao2017survey} has emerged as a promising technological solution to overcome the challenges faced by mobile devices (MDs) when performing high computational tasks, such as real-time data processing and artificial intelligence applications \cite{zhou2019edge} \cite{yousefpour2019all}. In spite of the MDs' technological advancements, their limited computing power and battery may lead to task drops, processing delays, and an overall poor user experience. By offloading intensive tasks to nearby edge nodes (ENs), MEC effectively empowers computation capability and reduces the delay and energy consumption. This improvement enhances the users' QoE, especially for time-sensitive computation tasks \cite{TNSE-QOE-24} \cite{ shah2018hierarchical}.

@@

\subsection{Related Work}
%Researchers have recently proposed several task offloading algorithms to tackle the aforementioned issues. Mao \textit{et al.} in \cite{mao2016dynamic} introduced a computation offloading algorithm for MEC systems. This scheme aims to reduce the MD's energy consumption while meeting the computation delay constraints. In \cite{zhang2016energy}, Zhang \textit{et al.} proposed an offloading scheme for MEC in heterogeneous cellular networks, taking into account the energy-constrained MEC scenario and the heterogeneity of MDs when optimizing the offloading decisions. Bi \textit{et al.} in \cite{bi2018computation} proposed an algorithm to optimize decision-making about computation offloading and power transferring in a wireless-powered MEC. Ning \textit{et al.} in \cite{ning2018cooperative} introduced a heuristic algorithm designed to enhance the efficiency of the partial computation offloading model. The works in \cite{mao2016dynamic}-\cite{ning2018cooperative} primarily focus on quasi-static systems and are not well-suited for dynamic systems and time-varying conditions.

@@

To cope with the dynamic nature of the network, recent research has proposed several task offloading algorithms using machine learning methods. In particular, deep reinforcement learning (DRL) hold promises to determine optimal decision-making policies by capturing the dynamics of environments and learning strategies for accomplishing long-term objectives \cite{arulkumaran2017deep}. DRL can effectively tackle the challenges of MEC arising from the ever-changing nature of networks, MDs, and servers' heterogeneity. This ultimately improves the MD users' QoE.

In \cite{huang2019deep}, Huang \textit{et al.} focused on a wireless-powered MEC. They proposed a DRL-based  approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered Internet of Things (IoT) devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility. In this approach, each MD requires quality-of-service information from other MDs. Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. 
%In \cite{chen2019energy}, Chen \textit{et al.} introduced an online algorithm with polynomial time complexity for offloading in MEC. This algorithm achieves a close-to-optimal energy consumption while ensuring that the delay is bounded. 
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm to concurrently optimize resource allocation and computation offloading. In \cite{huang2021deadline}, Huang \textit{et al.} proposed a DRL-based method based on a partially observable MDP, which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. Liu  \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. Zhou \textit{et al.} in \cite{zhou2021deep} used an MDP to study MEC and modeled the interactions of the environment. They proposed a Q-learning approach to achieve optimal resource allocation strategies and computation offloading. In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the industrial IoT systems to jointly optimize task offloading and resource allocation in order to achieve lower  energy consumption and decreased task delay. Liao \textit{et al.} in \cite{liao2023online} introduced a double reinforcement learning algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption.

\begin{table*}[htbp]
	\renewcommand{\arraystretch}{0.7}
	\captionsetup{name=TABLE}
	\caption{Comparison of Related Works}
	\scalebox{1}{%
		\begin{tabular}{ lp{1.8cm}p{3cm}p{3.3cm}p{0.8cm}p{2.2cm}p{2.5cm}l} 
			\toprule
			\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Algorithm/Methods} &  \textbf{Drawbacks}  \\ \midrule
			\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
			\cite{Bolourian-WCL24} & MEC with EH devices& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & Deep Q-Learning & \\\midrule
			\cite{zhao2019deep} & MEC & Wireless channels competition & Optimize downlink utility & MDP & Multi-agent D3QN & \\\midrule
			\cite{9253665} &  MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
			\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
			\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & \\\midrule
			\cite{huang2021deadline} & MEC & Computation offloading & Minimizing total energy consumption & POMDP & DDPG & \\ \midrule
			\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & MADRL & \\\midrule
			\cite{zhou2021deep} & MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule
			\cite{gao2022large} & heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent DDPG & \\\midrule
			\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & \\\midrule
			\cite{liao2023online} & MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDQN & \\\midrule
			\cite{wu2023multi} & MEC-based IIoT & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN& \\\midrule
			\cite{wu2023computation} & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & MARL & \\\midrule
			\cite{wu2024privacy} & MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent DDPG & \\\midrule
			\cite{wu2024combining} & MEC-based IIoT & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & DDPG  & \\\midrule
			QECO &  MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
			\toprule
	\end{tabular}}
	\label{table}
\end{table*}

@@

In this study, we delve into the computation task offloading problem in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We propose a distributed QoE-oriented computation offloading (QECO) algorithm that leverages DRL to efficiently handle task offloading in uncertain loads at ENs. This algorithm empowers MDs to make offloading decisions utilizing only locally observed information, such as task size, queue details, battery status, and historical workloads at the ENs. By adopting the appropriate policy based on each MD’s specific requirements at any given time, the QECO algorithm significantly improves the QoE for individual users.

@@

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QECO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's battery level, our approach leverages deep Q-network (DQN) \cite{mnih2015human} and long short-term memory (LSTM) \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors. We also analyze the training convergence and complexity of the proposed algorithm.

@@

where $p_i^{\text{C}} = 10^{-27}(f_i)^3$ represents the energy consumption of MD $i$'s CPU frequency \cite{mao2016dynamic}.

@@

We introduce $B_j(t) \triangleq |\mathcal{B}_j(t)|$ that represents the number of active queues in EN $j \in \mathcal{J}$ in time slot $t \in \mathcal{T}$. In each time slot $t \in \mathcal{T}$, the EN's processing power is divided among its active queues using a generalized processor sharing method~\cite{parekh1993generalized}. Let variable $f_j^{\text{E}}$ (in cycles per second) represent the computational capacity of EN $j$. Therefore, EN $j$ can allocate computational capacity of $f_j^{\text{E}}/(\rho_i(t) B_j(t))$ to each MD $i \in \mathcal{B}_j(t)$ during time slot $t$. To calculate the length of the computation queue for MD $i \in \mathcal{I}$ in EN $j \in \mathcal{J}$, we define $\omega_{i,j}(t)$ (in bits) to represent the number of bits from dropped tasks in that queue at the end of time slot $t \in \mathcal{T}$. The backlog of the queue, referred to as $\eta_{i,j}^{\text{E}}(t)$ is given by:

@@

	\eta_{i,j}^{\text{E}}(t)\hspace{-0.8mm}=\hspace{-1mm}\left[\eta_{i,j}^{\text{E}}(t-1)\hspace{-0.6mm}+\hspace{-0.7mm}\lambda_{i,j}^{\text{E}}(t)\hspace{-0.7mm}-\hspace{-0.7mm}{f_j^{\text{E}}\tau\over \rho_i(t)B_j(t)} -\omega_{i,j}(t)\right]^+\hspace{-1mm}.

@@

	\sum_{t^{'}=\hat{l}_{i,j}^{\text{E}}(t)}^{l_{i,j}^{\text{E}}(t)}{f_j^{\text{E}}\tau \over \rho_i(t)B_j(t^{'})}\mathbbm{1}(i \in \mathcal{B}_j(t^{'}))  \geq   \lambda_{i,j}^{\text{E}}(t),

@@

	\sum_{t^{'}=\hat{l}_{i,j}^{\text{E}}(t)}^{l_{i,j}^{\text{E}}(t)-1}{f_j^{\text{E}}\tau \over \rho_i(t)B_j(t^{'})}\mathbbm{1}(i \in \mathcal{B}_j(t^{'})) < \lambda_{i,j}^{\text{E}}(t),

@@

Additionally, we define $D_{i,j}^{\text{E}}(t)$ to represent the quantity of processing time slots allocated to task $z_{i,j}^{\text{E}}(t)$ when executed at EN $j$. This value is given by:
\begin{alignat}{1}
	D_{i,j}^{\text{E}}(t) = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over f_j^{\text{E}} \tau /  B_j(t)}.
	\label{14}  
\end{alignat}
We also define $E_{i,j}^{\text{E}}(t)$ as the energy consumption of processing at EN $j$ in time slot $t$ by MD $i$. This can be calculated as:
\begin{alignat}{1}
	E_{i,j}^{\text{E}}(t) = {D_{i,j}^{\text{E}}(t)  p_j^{\text{E}} \tau \over B_j(t)},  % = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over  f_j^{\text{E}}}p_j^{\text{E}} ,
	\label{15}  
\end{alignat}
where $p_j^{\text{E}}$ is a constant value which denotes the energy consumption of the EN $j$'s processor when operating at full capacity.

@@

Based on the introduced system model, we present the computation task offloading problem in this section. Our primary goal is to enhance each MD's QoE individually by taking the dynamic demands of MDs into account. To achieve this, we approach the optimization problem as an MDP, aiming to maximize the MD's QoE by striking a balance among key QoE factors, including task completion, task delay, and energy consumption. To prioritize QoE factors, we utilize the MD's battery level, which plays a crucial role in decision-making. Specifically, when an MD observes its state (e.g. task size, queue details, and battery level) and encounters a newly arrived task, it selects an appropriate action for that task. The selected action, based on the observed state, will result in enhanced QoE. Each MD strives to maximize its long-term QoE by optimizing the policy mapping from states to actions. In what follows, we first present the state space, action space, and QoE function, respectively. We then formulate the QoE maximization problem for each MD.

\subsection{State Space}
A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state in time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's battery level, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
\begin{alignat}{1}
\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
	\label{18}
\end{alignat}
where vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1))_{j \in \mathcal{J}}$ represents the queues length of MD $i$ in ENs at the previous time slot and is computed by the MD according to \eqref{10}. Let $\phi_i(t)$ denote the battery level of MD $i$ in time slot $t$. Considering the power modes of a real mobile device, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

In addition, to predict future EN workloads, we define the matrix $\mathcal{H}(t)$ as historical data, indicating the number of active queues for all ENs. This data is recorded over $T^{\text{s}}$ time slots, ranging from $t\hspace{-0.4mm}-\hspace{-0.4mm}T^{\text{s}}$ to $t\hspace{-0.4mm}-\hspace{-0.4mm}1$, in $T^{\text{s}}\hspace{-0.4mm}\times \hspace{-0.4mm} J$ matrix. For EN $j$ workload history at $i^{th}$ time slot from $T^{\text{s}}-t$, we define $h_{i,j}(t)$ as:
\begin{alignat}{1}
	h_{i,j}(t) = B_j(t - T^{\text{s}} + i - 1).
	\label{19}
\end{alignat}
%where $B_j(t - T^{\text{s}} + i - 1)$ is the number of active queues of EN $j$ in time slot $t - T^{\text{s}} + i - 1$. 
EN $j \in \mathcal{J}$ broadcasts $B_j(t)$ at the end of each time slot. %Given $B_j(t) \leq I$, can be represented using only a few bits, the broadcast incurs minimal signaling overhead.

@@

$\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
\begin{alignat}{1}
\phi_i(t) \, \mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) +(1-\phi_i(t)) \, \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)),
	\label{23}  
\end{alignat}
where $\phi_i(t)$ represents the MD $i$'s battery level. When the MD is operating in performance mode, the primary focus is on minimizing task delays, thus the delay contributes more to the cost. On the other hand, when the MD switches to ultra power-saving mode, the main attention is directed toward reducing power consumption.

>>>>

		\lettrine{M}{ \,OBILE} edge computing (MEC) \cite{mao2017survey} has emerged as a promising technological solution to overcome the challenges faced by mobile devices (MDs) when performing high computational tasks, such as real-time data processing and artificial intelligence applications \cite{zhou2019edge} \cite{yousefpour2019all}. In spite of the MDs' technological advancements, their limited computing power and \textcolor{blue}{energy} may lead to task drops, processing delays, and an overall poor user experience. By offloading intensive tasks to nearby edge nodes (ENs), MEC effectively empowers computation capability and reduces the delay and energy consumption. This improvement enhances the users' QoE, especially for time-sensitive computation tasks \cite{TNSE-QOE-24} \cite{ shah2018hierarchical}.

@@

		%methods, advances, drawbacks \textcolor{blue}
		\begin{table*}[htbp]\textcolor{blue}{
				\renewcommand{\arraystretch}{0.7}
				\captionsetup{name=TABLE}
				\caption{Comparison of Related Works}
				\scalebox{1}{%
					\begin{tabular}{ lp{1.7cm}p{3cm}p{3.4cm}p{0.8cm}p{2.2cm}p{3cm}l} 
						\toprule
						\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Algorithm/Methods} &  \textbf{Drawbacks}  \\ \midrule
						\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
						\cite{Bolourian-WCL24} & MEC with EH devices& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & Deep Q-Learning & Not consider computation delay  \\\midrule
						\cite{zhao2019deep} & MEC & User association and resource allocation & Optimize downlink utility & MDP & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
						\cite{9253665} &  MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
						\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
						\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & Not take user's deman into consideration \\\midrule
						\cite{huang2021deadline} & MEC & Computation offloading & Optimize drop rate and energy consumption & POMDP & DDPG & Not consider computation delay \\ \midrule
						\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & MADRL &  Not consider delay-sensitive tasks \\\midrule
						\cite{zhou2021deep} & heterogeneous MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule
						\cite{gao2022large} & heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent DDPG & Not take user's deman into consideration \\\midrule
						\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & Not take user's deman into consideration\\\midrule
						\cite{liao2023online} & MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDQN & Only for a singel MEC system model  \\\midrule
						\cite{wu2023multi} & MEC-based IIoT & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN& \\\midrule
						\cite{wu2023computation} & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & MARL & \\\midrule
						\cite{wu2024privacy} & MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent DDPG & \\\midrule
						\cite{wu2024combining} & MEC-based IIoT & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & DDPG  & \\\midrule
						QECO &  MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
						\toprule
				\end{tabular}}
				\label{table}}
		\end{table*}

		\subsection{Related Work}

		%Researchers have recently proposed several task offloading algorithms to tackle the aforementioned issues. Mao \textit{et al.} in \cite{mao2016dynamic} introduced a computation offloading algorithm for MEC systems. This scheme aims to reduce the MD's energy consumption while meeting the computation delay constraints. In \cite{zhang2016energy}, Zhang \textit{et al.} proposed an offloading scheme for MEC in heterogeneous cellular networks, taking into account the energy-constrained MEC scenario and the heterogeneity of MDs when optimizing the offloading decisions. Bi \textit{et al.} in \cite{bi2018computation} proposed an algorithm to optimize decision-making about computation offloading and power transferring in a wireless-powered MEC. Ning \textit{et al.} in \cite{ning2018cooperative} introduced a heuristic algorithm designed to enhance the efficiency of the partial computation offloading model. The works in \cite{mao2016dynamic}-\cite{ning2018cooperative} primarily focus on quasi-static systems and are not well-suited for dynamic systems and time-varying conditions.

@@

		wireless channels competition  optimize downlink utility

		To cope with the dynamic nature of the network, recent research has proposed several task offloading algorithms using machine learning methods. In particular, deep reinforcement learning (DRL) hold promises to determine optimal decision-making policies by capturing the dynamics of environments and learning strategies for accomplishing long-term objectives \cite{arulkumaran2017deep}. DRL can effectively tackle the challenges of MEC arising from the ever-changing nature of networks, MDs, and servers' heterogeneity. This ultimately improves the MD users' QoE. In \cite{huang2019deep}, Huang \textit{et al.} focused on a wireless-powered MEC. They proposed a DRL-based  approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered Internet of Things (IoT) devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility. In this approach, each MD requires quality-of-service information from other MDs. Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. 
		%In \cite{chen2019energy}, Chen \textit{et al.} introduced an online algorithm with polynomial time complexity for offloading in MEC. This algorithm achieves a close-to-optimal energy consumption while ensuring that the delay is bounded. 
		Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm to concurrently optimize resource allocation and computation offloading. In \cite{huang2021deadline}, Huang \textit{et al.} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. Liu  \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL (MADRL), which can generate interactive information along with resource decisions. Zhou \textit{et al.} in \cite{zhou2021deep} used an MDP to study MEC and modeled the interactions of the environment. They proposed a Q-learning approach to achieve optimal resource allocation strategies and computation offloading. In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the industrial IoT systems to jointly optimize task offloading and resource allocation in order to achieve lower  energy consumption and decreased task delay. Liao \textit{et al.} in \cite{liao2023online} introduced a double reinforcement learning algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption.

@@

		In this study, we delve into the computation task offloading problem in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We propose a distributed QoE-oriented computation offloading (QECO) algorithm that leverages DRL to efficiently handle task offloading in uncertain loads at ENs. This algorithm empowers MDs to make offloading decisions utilizing only locally observed information, such as task size, queue details, \textcolor{blue}{energy status}, and historical workloads at the ENs. By adopting the appropriate policy based on each MD’s specific requirements at any given time, the QECO algorithm significantly improves the QoE for individual users.

@@

			\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QECO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's \textcolor{blue}{energy level}, our approach leverages deep Q-network (DQN) \cite{mnih2015human} and long short-term memory (LSTM) \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors. We also analyze the training convergence and complexity of the proposed algorithm.

@@

		where $p_i^{\text{C}} = 10^{-27}(f_i)^3$ represents the energy consumption of MD $i$'s CPU frequency []. %\cite{mao2016dynamic}.

@@

		We introduce $b_j(t) \triangleq |\mathcal{B}_j(t)|$ that represents the number of active queues in EN $j \in \mathcal{J}$ in time slot $t \in \mathcal{T}$. In each time slot $t \in \mathcal{T}$, the EN's processing power is divided among its active queues using a generalized processor sharing method~\cite{parekh1993generalized}. Let variable $f_j^{\text{E}}$ (in cycles per second) represent the computational capacity of EN $j$. Therefore, EN $j$ can allocate computational capacity of $f_j^{\text{E}}/(\rho_i(t) b_j(t))$ to each MD $i \in \mathcal{B}_j(t)$ during time slot $t$. To calculate the length of the computation queue for MD $i \in \mathcal{I}$ in EN $j \in \mathcal{J}$, we define $\omega_{i,j}(t)$ (in bits) to represent the number of bits from dropped tasks in that queue at the end of time slot $t \in \mathcal{T}$. The backlog of the queue, referred to as $\eta_{i,j}^{\text{E}}(t)$ is given by:

@@

			\eta_{i,j}^{\text{E}}(t)\hspace{-0.8mm}=\hspace{-1mm}\left[\eta_{i,j}^{\text{E}}(t-1)\hspace{-0.6mm}+\hspace{-0.7mm}\lambda_{i,j}^{\text{E}}(t)\hspace{-0.7mm}-\hspace{-0.7mm}{f_j^{\text{E}}\tau\over \rho_i(t)b_j(t)} -\omega_{i,j}(t)\right]^+\hspace{-1mm}.

@@

			\sum_{t^{'}=\hat{l}_{i,j}^{\text{E}}(t)}^{l_{i,j}^{\text{E}}(t)}{f_j^{\text{E}}\tau \over \rho_i(t)b_j(t^{'})}\mathbbm{1}(i \in \mathcal{B}_j(t^{'}))  \geq   \lambda_{i,j}^{\text{E}}(t),

@@

			\sum_{t^{'}=\hat{l}_{i,j}^{\text{E}}(t)}^{l_{i,j}^{\text{E}}(t)-1}{f_j^{\text{E}}\tau \over \rho_i(t)b_j(t^{'})}\mathbbm{1}(i \in \mathcal{B}_j(t^{'})) < \lambda_{i,j}^{\text{E}}(t),

@@

		Additionally, we define $D_{i,j}^{\text{E}}(t)$ to represent the quantity of processing time slots allocated to task $z_{i,j}^{\text{E}}(t)$ when executed at EN $j$. This value is given by:
		\begin{alignat}{1}
			D_{i,j}^{\text{E}}(t) = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over f_j^{\text{E}} \tau /  b_j(t)}.
			\label{14}  
		\end{alignat}
		We also define $E_{i,j}^{\text{E}}(t)$ as the energy consumption of processing at EN $j$ in time slot $t$ by MD $i$. This can be calculated as:
		\begin{alignat}{1}
			E_{i,j}^{\text{E}}(t) = {D_{i,j}^{\text{E}}(t)  p_j^{\text{E}} \tau \over b_j(t)},  % = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over  f_j^{\text{E}}}p_j^{\text{E}} ,
			\label{15}  
		\end{alignat}
		where $p_j^{\text{E}}$ is a constant value which denotes the energy consumption of the EN $j$'s processor when operating at full capacity.

@@

		Based on the introduced system model, we present the computation task offloading problem in this section. Our primary goal is to enhance each MD's QoE individually by taking the dynamic demands of MDs into account. To achieve this, we approach the optimization problem as an MDP, aiming to maximize the MD's QoE by striking a balance among key QoE factors, including task completion, task delay, and energy consumption. To prioritize QoE factors, we utilize the MD's \textcolor{blue}{energy level}, which plays a crucial role in decision-making. Specifically, when an MD observes its state (e.g. task size, queue details, and \textcolor{blue}{energy status}) and encounters a newly arrived task, it selects an appropriate action for that task. The selected action, based on the observed state, will result in enhanced QoE. Each MD strives to maximize its long-term QoE by optimizing the policy mapping from states to actions. In what follows, we first present the state space, action space, and QoE function, respectively. We then formulate the QoE maximization problem for each MD.

		\subsection{State Space}
		A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state in time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's \textcolor{blue}{energy level}, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
		\begin{alignat}{1}
			\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
			\label{18}
		\end{alignat}
		where vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1))_{j \in \mathcal{J}}$ represents the queues length of MD $i$ in ENs at the previous time slot and is computed by the MD according to \eqref{10}. Let $\phi_i(t)$ denote the \textcolor{blue}{energy level} of MD $i$ in time slot $t$. Considering the power modes of \textcolor{blue}{some} real mobile device, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

		In addition, to predict future EN workloads, we define the matrix $\mathcal{H}(t)$ as historical data, indicating the number of active queues for all ENs. This data is recorded over $T^{\text{s}}$ time slots, ranging from $t\hspace{-0.4mm}-\hspace{-0.4mm}T^{\text{s}}$ to $t\hspace{-0.4mm}-\hspace{-0.4mm}1$, in $T^{\text{s}}\hspace{-0.4mm}\times \hspace{-0.4mm} J$ matrix. For EN $j$ workload history at $i^{th}$ time slot from $T^{\text{s}}-t$, we define $h_{i,j}(t)$ as:
		\begin{alignat}{1}
			h_{i,j}(t) = b_j(t - T^{\text{s}} + i - 1).
			\label{19}
		\end{alignat}
		%where $b_j(t - T^{\text{s}} + i - 1)$ is the number of active queues of EN $j$ in time slot $t - T^{\text{s}} + i - 1$. 
		EN $j \in \mathcal{J}$ broadcasts $b_j(t)$ at the end of each time slot. %Given $b_j(t) \leq I$, can be represented using only a few bits, the broadcast incurs minimal signaling overhead.

@@

		$\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
		\begin{alignat}{1}
			\phi_i(t) \, \mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) +(1-\phi_i(t)) \, \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)),
			\label{23}  
		\end{alignat}
		where $\phi_i(t)$ represents the MD $i$'s \textcolor{blue}{energy level}. When the MD is operating in performance mode, the primary focus is on minimizing task delays, thus the delay contributes more to the cost. On the other hand, when the MD switches to ultra power-saving mode, the main attention is directed toward reducing power consumption.
